---
title: "Lab 5"
author: "Shayan Kaveh"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Confidence intervals

In this section, we'll learn to use the 't.test()' function as a way to return the confidence interval using the t-distribution.

## Example 1

Create a mock sample of observation values and find the confidence interval. Then we'll learn how to report that.

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 102)

pika_mean <- mean(pika_mass) # calculate mean

pika_sd <- sd(pika_mass) # calculate standard deviation

pika_ci <- t.test(pika_mass) # this exists in base R, so it'll work without any attached packages

pika_ci
```

### Risky & tedious way:

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g.

### A high initial investment, but safer & better way:

In-line code referencing to stored objects!

Use a single back tick (*not* an apostrophe) on either side of a lowercase r to create an in-line bit of code.

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r round(pika_ci$conf.int[1], 1)`, `r round(pika_ci$conf.int[2], 1)`] g.

beautiful!!! looks exactly the same in the html, but if I change any values, they'll be updated automatically!

# Two-sample t-test

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0).

- H0: Means difference = 0
- HA: Means difference is NOT = 0

This is a two-sided t-test because no directionality is implied.

## Example 1

```{r}
oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7)
  
pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)
```

Is there a significant difference in mean heights for oak and pine trees?

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

The p-value of `r trees_t$p.value` means that there is a `r round(trees_t$p.value * 100, 1)`% chance of finding sample means *at least as different* as those I found by random chance if they were drawn from populations with the same mean height.

Based on this, I would reject the null hypothesis. 

# Now let's use real data

Using the `mpg` dataset to compare city gas mileage for SUVs and compact cars.

```{r}
# create a subset called compact_suv that only contains observations where the class is SUV or compact

compact_suv <- mpg %>% 
  filter(class %in% c("suv", "compact"))
```

# Create a histogram and a QQ-plot
```{r}
ggplot(data = compact_suv, 
       aes(x = cty)) +
  geom_histogram(aes(fill = class, color = class)) +
  facet_wrap(~ class) +
  theme_classic()

ggplot(data = compact_suv,
       aes(sample = cty)) + 
  geom_qq() +
  facet_wrap(~ class) + 
  theme_classic()

# The result looks fairly linear for both other than a few values at the tail ends. This means the means are normally distributed.
```

```{r}
# Let's make a summary table to look at the statistical values!

car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(mean_city = round(mean(cty), 1),
            sd_city = round(sd(cty), 1),
            median_city = median(cty),
            sample_size = n())

car_stats
```

```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty) # this creates a vector of values rather than a new data source
  
suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)
```

stats speaks: I reject the null hypothesis of equal mean city gas mileages

Mean gas mileage for compact cars and SUVs differs significantly (t(`r round(cars_t$parameter, 1)` = `r round(cars_t$statistic, 1)`, p < 0.001, $\alpha$ = 0.05))
