---
title: "Lab 5"
author: "Shayan Kaveh"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Confidence intervals

In this section, we'll learn to use the 't.test()' function as a way to return the confidence interval using the t-distribution.

## Example 1

Create a mock sample of observation values and find the confidence interval. Then we'll learn how to report that.

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 102)

pika_mean <- mean(pika_mass) # calculate mean

pika_sd <- sd(pika_mass) # calculate standard deviation

pika_ci <- t.test(pika_mass) # this exists in base R, so it'll work without any attached packages

pika_ci
```

### Risky & tedious way:

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g.

### A high initial investment, but safer & better way:

In-line code referencing to stored objects!

Use a single back tick (*not* an apostrophe) on either side of a lowercase r to create an in-line bit of code.

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r round(pika_ci$conf.int[1], 1)`, `r round(pika_ci$conf.int[2], 1)`] g.

beautiful!!! looks exactly the same in the html, but if I change any values, they'll be updated automatically!

# Two-sample t-test

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0).

- H0: Means difference = 0
- HA: Means difference is NOT = 0

This is a two-sided t-test because no directionality is implied.

## Example 1

```{r}
oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7)
  
pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)
```

Is there a significant difference in mean heights for oak and pine trees?

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

The p-value of `r trees_t$p.value` means that there is a `r round(trees_t$p.value * 100, 1)`% chance of finding sample means *at least as different* as those I found by random chance if they were drawn from populations with the same mean height.

Based on this, I would reject the null hypothesis. 
